---
title: "Eudora's BGM analysis"
output:
  html_document: 
    keep_md: TRUE
editor_options: 
  chunk_output_type: console
---

# Importing data and pre-processing

```{r, warning=FALSE, error=FALSE, message=FALSE}
library(phyloseq)
library(ggplot2)
library(tidyverse)
library(RColorBrewer)
library(cowplot)
library(metacoder)

BGM_data0 <- import_biom(BIOMfilename = "otu_table_fungi.biom")
BGM_data0
```

## Checking sequencing depth

```{r}
# Make a data frame with a column for the read counts of each sample
sample_sum_df <- data.frame(sum = sample_sums(BGM_data0))

# Histogram of sample read counts
ggplot(sample_sum_df, aes(x = sum)) + 
  geom_histogram(color = "black", fill = "indianred", binwidth = 2500) +
  ggtitle("Distribution of sample sequencing depth") + 
  xlab("Read counts") +
  theme_gray() + theme(axis.title.y = element_blank())
```

## Standardizing by sequencing depth

```{r}
#Standardize abundances to the median sequencing depth
total <- median(sample_sums(BGM_data0))
standf <- function(x, t=total) round(t * (x/sum(x)))

BGM_data0.std <- transform_sample_counts(BGM_data0, standf)

#Filter taxa with cutoff 3.0 Coefficient of Variation
#BGM_data0.stdf <- filter_taxa(BGM_data0.std, function(x) sd(x)/mean(x) > 3.0, TRUE)
```

## Filtering "no hits"

```{r}
#Renaming taxonomy levels on tax_table
colnames(tax_table(BGM_data0.std)) <- c("Kingdom", "Phylum", "Class","Order", "Family", "Genus", "Species")

#Filtering no hit at Kingdom level
BGM_data <- subset_taxa(BGM_data0.std, Kingdom != "No blast hit")
```

# Taxa abudance at different rank levels

```{r, fig.width= 10, fig.height=8, fig.align='center'}
#summarizing by tax rank
BGM.order <- tax_glom(BGM_data, "Phylum")

plot_bar(BGM_data, fill = "Phylum", x = "Group")

plot_bar(BGM.order, fill = "Phylum", x = "Group")

#Plot by family
BGM.fam <- tax_glom(BGM_data, "Class")
plot_bar(BGM.fam, fill = "Class", x = "Group")

```

## Subsetting datasets by study
```{r}
#Silvehill mine dataset
SH_data <- subset_samples(BGM_data, grepl("SH", sample_data(BGM_data)$Group))


#BGM mine dataset
BGM_data.f <- subset_samples(BGM_data, !grepl("SH", sample_data(BGM_data)$Group))
```

# Brewer Gold Mine study

## Relative abudance of genus and class in BGM soil and root samples

```{r, fig.width= 10, fig.height=8, fig.align='center'}
TopNOTUs <- function(sample,N) {
  names(sort(taxa_sums(sample), TRUE)[1:N])
}


#Creating data frame from phyloseq object
top.BGM <- TopNOTUs(BGM_data.f, 50)
BGM.df <- prune_taxa(top.BGM, BGM_data.f) %>% psmelt()

#Sumarize data by site 
#test <- BGM.df %>% group_by(Genus) %>% summarise(sum(Abundance))

#plot by relative abundance
#Plot
pal <- colorRampPalette((brewer.pal(10, "Paired")))

Genus_soil <- ggplot(data = filter(BGM.df, Description == "Soil"), aes(Group, Abundance, fill = Genus)) +
  geom_bar(stat = "identity", position = position_fill()) + coord_flip() +
    scale_fill_manual(values = (rev(pal(24)))) + 
    guides(fill = guide_legend(reverse = TRUE, ncol = 1, keyheight = 0.8)) +
    facet_grid(~ Description, drop = TRUE) +
   theme(text = element_text(size = 15)) + theme_gray()


Genus_root <- ggplot(data = filter(BGM.df, Description == "Root"), aes(Group, Abundance, fill = Genus)) +
  geom_bar(stat = "identity", position = position_fill()) + coord_flip() +
    scale_fill_manual(values = rev(pal(24))) + 
    guides(fill = FALSE) +
    facet_grid(~ Description, drop = TRUE) +
   theme(text = element_text(size = 15)) + theme_gray()

plot_grid(Genus_root, Genus_soil, align = "h", rel_widths = c(1,1.2))

Class_soil <-ggplot(data = filter(BGM.df, Description == "Soil"), aes(Group, Abundance, fill = Class)) +
  geom_bar(stat = "identity", position = position_fill()) + coord_flip() +
  scale_fill_manual(values = (pal(24))) + 
  guides(fill = guide_legend(ncol = 1, keyheight = 0.8)) +
  facet_grid(~ Description, drop = TRUE) +
  theme(text = element_text(size = 15)) + theme_gray()

Class_root <-ggplot(data = filter(BGM.df, Description == "Root"), aes(Group, Abundance, fill = Class)) +
  geom_bar(stat = "identity", position = position_fill()) + coord_flip() +
  scale_fill_manual(values = (pal(24))) + 
  guides(fill = FALSE) +  
  facet_grid(~ Description, drop = TRUE) +
  theme(text = element_text(size = 15)) + theme_gray()

plot_grid(Class_root, Class_soil, align = "h", rel_widths = c(1,1.6))
```

## Metacoder analysis for BGM soil and root samples

```{r, warning=FALSE, message=FALSE}
#New variable
sample_data(BGM_data.f)$site_type <- str_sub(sample_data(BGM_data.f)$Site, 1,2)

#Top 100 OTUs

top100.BGM <- TopNOTUs(BGM_data.f, 200)
top100.BGM.ps <- prune_taxa(top100.BGM, BGM_data.f)


#Converting to metacoder
obj <- parse_phyloseq(top100.BGM.ps)

# Convert counts to proportions
obj$data$otu_table <- calc_obs_props(obj,
                                     data = "otu_table",
                                     cols = obj$data$sample_data$sample_id)
# Calculate per-taxon proportions
obj$data$tax_table <- calc_taxon_abund(obj,
                                       data = "otu_table",
                                       cols = obj$data$sample_data$sample_id)

#Compare treatments
obj$data$diff_table <- compare_groups(obj,
                                          data = "tax_table",
                                          cols = obj$data$sample_data$sample_id,
                                          groups = obj$data$sample_data$site_type)
```

```{r, fig.width= 10, fig.height=8, fig.align='center'}
#Tree visual
set.seed(1)
Tree1 <- metacoder::heat_tree(taxa::filter_taxa(obj, taxon_names == "c__Agaricomycetes", subtaxa = TRUE),
          node_size = n_obs, 
          node_label = taxon_names,
          node_color = log2_median_ratio,
          node_color_range = c("#a6611a","#dfc27d","#bdbdbd","#80cdc1","#018571"), 
          node_color_trans = "linear",
          node_label_max = 120,
          node_color_interval = c(-1, 1),
          edge_color_interval = c(-1, 1),
          node_size_axis_label = "Number of OTUs",
          node_color_axis_label = "Log2 ratio median proportions",
          title = "Contaminated site vs Adjacent Forest",
          title_size = 0.03,
          initial_layout = "reingold-tilford", layout = "davidson-harel")

Tree1
```

## Funguild analysis of BGM soil and root samples

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#Reading function
source("fg_assign.r")

str_charc <- function(x) {gsub('[f,p,c,o,k,g,s]__', '', x)}

#Modifying table to assign guild
test <- data.frame(tax_table(BGM_data.f)) %>%
  dplyr::rename(kingdom = Kingdom, 
                phylum = Phylum, 
                class = Class, 
                order = Order, 
                family = Family, 
                genus = Genus, 
                species = Species) %>% 
  rownames_to_column( var = "OTU_ID") %>%
  mutate_all(.funs = str_charc) %>%
  column_to_rownames("OTU_ID")
                    
#Assigning guild
test.assign <- fg_assign(test) %>%
  rownames_to_column(var = "OTU")
```

```{r, warning=FALSE, message=FALSE}
#Assigning guild to the tax table
new_tax <- data.frame(tax_table(BGM_data.f)) %>% 
  rownames_to_column(var = "OTU") %>% 
  left_join(test.assign[,c("OTU","guild")], by = "OTU") %>% 
  select(OTU, Kingdom, Phylum, Class, Order, Family, guild, Genus, Species) %>% 
  column_to_rownames(var = "OTU")
  
new_tax.0 <- tax_table(as.matrix(new_tax))
tax_table(BGM_data.f) <- new_tax.0
```

```{r, fig.width= 10, fig.height=8, fig.align='center'}
#Plotting abudance by guild
#Creating data frame from phyloseq object
top.BGM <- TopNOTUs(BGM_data.f, 50)
BGM.df <- prune_taxa(top.BGM, BGM_data.f) %>% psmelt()

#Plot
pal <- colorRampPalette((brewer.pal(8, "Paired")))

Guild_soil <- ggplot(data = filter(BGM.df, Description == "Soil" & !is.na(guild)), aes(Group, Abundance, fill = guild)) +
  geom_bar(stat = "identity", position = position_fill()) + coord_flip() +
  scale_fill_manual(values = pal(16)) + 
  guides(fill = guide_legend(ncol = 3, keyheight = 0.6)) +
  facet_grid(~ Description, drop = TRUE) +
  theme_gray()

Guild_soil.0 <- Guild_soil + theme(legend.position = "none")
Guild_soil.1 <- Guild_soil + theme(legend.position = "bottom")

Guild_root <- ggplot(data = filter(BGM.df, Description == "Root" & !is.na(guild)), aes(Group, Abundance, fill = guild)) +
  geom_bar(stat = "identity", position = position_fill()) + coord_flip() +
  scale_fill_manual(values = pal(16)) + 
  guides(fill = FALSE) +  
  facet_grid(~ Description, drop = TRUE) +
  theme(text = element_text(size = 15)) + theme_gray()

P <- plot_grid(Guild_root, Guild_soil.0, align = "h", rel_heights = c(0.7,0.7))
ggdraw() + draw_plot(P, 0, 0.2, 1, 0.8) + draw_plot(legend, 0, -0.35, 1, 1, 2)
```

